---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
This is preliminary exploratory shit for the opioid project
```{r}
setwd("C:/Users/zgord/OneDrive - University of North Carolina at Chapel Hill/Semesters/Spring 2021/STOR 320/Project/Data")

# import libraries
library(tidyverse)
```


```{r}
# read in data
opioid_long = read.csv("Master_Table_Long.csv")
opioid_wide = read.csv("Wide_Master.csv")
opioid_wide %>% head()
opioid_wide %>% str()

```
Exploring population density.

```{r}
# I am selecting the variables I want to use to analyze population density. I also add a variable for Population density called `PopDensity` and a variable for Porportion dead, called `PorpDead`. 
opioid_pop =opioid_wide %>% 
  select(State,County,Year,Population,LandArea,Total) %>% 
  # LandArea is measured in square miles so PopDensity represents number of people per square mile.
  mutate(PopDensity = Population/LandArea,
         PorpDead = Total/Population,) %>% 
  mutate_all(~replace(., is.na(.),0))

```


```{r}
# This Chunk of code graphs change in population density over time. It only looks at years after 2011, because that where most of the values are. I then color it by state_county.
opioid_pop %>%
  group_by(State,County,Year) %>% 
  unite(state_county,State,County, remove=FALSE) %>% 
  filter(Year>=2011) %>% 
  ggplot(aes(x=Year,y=PopDensity, color = state_county)) +
  geom_line() +
  theme(legend.position = "none") +
  scale_y_log10()


```


Analasys: It is a bit hard to tell how population density changes over time. For counties with high population density it looks like there is a slight increase. In the orginal graph it was impossible to see what was happening with low population density areas. So I graphed all counties with populations below 500 person per square. It didn't really help. Looking at change in population over time was not useful.
<br>
<br>

This next graph looks at boxplots of the total number of opioid deaths by year and the number of people, per 100,000 dead by year.
```{r}
# This chunk of code looks at two box plots. The first is the total number of deaths by year.
opioid_pop %>% 
  ggplot(aes(x=Year %>% as.factor(),y=Total, group = Year %>% as.factor())) +
  geom_boxplot()

# This graph looks at porportion dead by year
opioid_pop %>% 
  ggplot(aes(x=Year %>% as.factor(),y=PorpDead, group=Year)) +
  geom_boxplot()
```
This doesn't give me a lot of data, because it doesn not distinguish between high density areas and low density areas.

The next Graph I want to look at will compare population density to total number of opioid deaths. I will also break it down by year because that seems important.
```{r fig.height=50, fig.width = 10}

opioid_pop %>%
  group_by(State,County,Year) %>% 
  unite(state_county,State,County, remove=FALSE) %>% 
  filter(Year>=2011) %>% 
  
    ggplot(aes(x=PopDensity,y=(Total/Population)*100000, group = state_county, color = State)) +
    geom_point() +
    theme(legend.position = "none")+
    scale_x_log10() +
    facet_wrap(~Year, nrow = 7)
```
This Chart indicates that as Population density increases the per capita deaths decrease. This is the graph I will use for this question in the template.

```{r }
opioid_wide
opioid_nub = opioid_wide %>% 
  select(State, County, Non_US_Born, Population, Year, Total) %>% 
  mutate(porp_nub = Non_US_Born/Population) %>% 
  group_by(Year) %>% 
  mutate(nub_quantile = ntile(porp_nub,20))

opioid_nub %>%  head()

opioid_nub %>% 
  ggplot(aes(y=porp_nub)) +
  geom_boxplot()


```


I will now look at the relationship between residency status and  
```{r fig.height=50, fig.width = 15}
opioid_nub %>% 
  ggplot(aes(x = nub_quantile %>% as.factor(), y = (Total/Population)*100000, group = nub_quantile)) +
  geom_boxplot() +
  facet_wrap(~Year, nrow = 7)
```


```{r include=FALSE}
library(leaflet)
library(maps)
```



I am experimenting with creating a map of opioid deaths with leaflet. It is very incomplete, all of the values are in the incorrect county. This is the first map I have made with leaflet so I am still exploring it.
```{r}
opioid_pop_2011 = opioid_pop %>% filter(Year == 2011)


pal = colorNumeric(
  palette = "Blues",
  domain = opioid_pop_2011$PorpDead*100000
)


labels <- sprintf(
  "<strong>%s, %s</strong><br/>%g opioid deaths / 100,000",
  opioid_pop_2011$County,opioid_pop_2011$State, opioid_pop_2011$PorpDead*100000
) %>% lapply(htmltools::HTML)



map("county", fill = T, plot=F) %>% 
  leaflet() %>% 
  addProviderTiles("CartoDB.Positron", options = providerTileOptions(opacity = 1)) %>% 
  addPolygons(stroke = T,
              smoothFactor = .2,
              color = "gray",
              weight = 2,
              fillOpacity =.5,
              fillColor = ~pal(opioid_pop_2011$PorpDead*100000),
              dashArray = 3,
              highlight = highlightOptions(
                weight = 3,
                fillOpacity = 1,
                color = "black",
                opacity = 1.0,
                bringToFront = TRUE),
              label = labels,
              labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto")) %>% 
  addLegend("bottomright",
            pal = pal,
            values = ~opioid_pop_2011$PorpDead*100000,
            title = "Opioid Deaths per 100,000")





```




























































